---
title: "Leeds Tree-planting Prioritization"
author: "Mark Miller"
date: "9 October 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries and read in attributed planning units

```{r}
library(dplyr)
library(sf)
library(prioritizr)
library(ggplot2)
#library(stargazer)

pu_all<-read_sf('C:/trees/modelling/demo/final_PU_attrib.shp')

```

Plot land cover types inside Planning Units, which form basis of cost layer. All land  Planning Unit landcover is theoretically available for planting and binned into 4 cost c lasses: existing tree cover = 0 (free); all others are 1,2 or 3 representing increasing difficulty/cost to plant on. 
```{r, echo=FALSE}

tab1<-pu_all%>%st_set_geometry(NULL)%>%group_by(prmryFn)%>%summarise(cost=first(cost))
tab1%>%as.data.frame()
#stargazer(tab1, type='text')
plot(pu_all['prmryFn'],border=NA)
#ggplot(pu_all)+geom_sf(aes(fill=prmryFn))+theme_bw()+ggtitle('Landcover')
```

Plot cost layer and other planning layers

```{r, echo=FALSE}
#ggplot(pu_all)+geom_sf(aes(fill=factor(cost)))+theme_bw()+ggtitle('Cost')
plot(pu_all['cost'],border=NA)

#ggplot(pu_all)+geom_sf(aes(fill=mst30dp))+theme_bw()+ggtitle('Highest 30% Deprivation')
plot(pu_all['mst30dp'],border=NA)
#ggplot(pu_all)+geom_sf(aes(fill=md30t60))+theme_bw()+ggtitle('Intermediate Deprivation')
plot(pu_all['md30t60'],border=NA)
#ggplot(pu_all)+geom_sf(aes(fill=flood))+theme_bw()+ggtitle('Priority flood planting')
plot(pu_all['flood'],border=NA)

plot(pu_all['car_pot'],border=NA)
#ggplot(pu_all)+geom_sf(aes(fill=car_pot))+theme_bw()+ggtitle('Carbon potential')
```

Initiate prioritizr problem and set up target parameters

```{r}
#convert area m2 to km2
#pu_all$flood<-pu_all$flood/1000000
#pu_all$car_pot<-pu_all$car_pot/1000000
#pu_all$area<-NULL

p1<-problem(as(pu_all, 'Spatial'), features=c('car_pot', 'mst30dp', 
             "md30t60", 'flood'),cost_column='cost')
```
We can have a max target of 39444 Planning Units that have high deprivation
```{r}
sum(pu_all$mst30dp, na.rm=T)
```
We can have a max target of 20635 Planning Units that have intermediate deprivation
```{r}
sum(pu_all$md30t60, na.rm=T)
```
We can have a max target of 51218396 m2 (~51 km2) area of priority flood planting
```{r}
sum(pu_all$flood, na.rm=T)
```
There is 365115467 m2 (~365 km2) area available for tree planting (maximum carbon potential)
```{r}
sum(pu_all$car_pot)
```
But 49490374 m2 (~49 km2) is already covered by trees, so the remaining area we could plant is 315625093 m2 (~315 km2). Any cabon potential target needs to be above 49 km2 otherwise we're not planting new trees
```{r}
sum(pu_all$car_pot)
sum(pu_all$car_pot)-sum(pu_all[pu_all$cost==0,]$car_pot, na.rm=T)
```

OK so set prelimary targets to double current tree cover (basically 100km2), include 50% of high deprivation planning units and 50% of intermediate deprivation planning units, and include half of avilable priority tree planting?
```{r}
targets<-c(100000000, 19722,10317, 25609198) 
```

Run prioritizr
```{r}
p2<-p1%>%add_min_set_objective() %>%   
  add_absolute_targets(targets)
#%>%  add_boundary_penalties(10, 1) # selects bigger blocks
presolve_check(p2)

s1<-solve(p2) # calls Gurobi
```

plot solution

```{r}
pu_all$sol<-s1$solution_1

#ggplot(pu_all)+geom_sf(aes(fill=factor(sol)))+theme_bw()+ggtitle('Planning solution 1')
plot(pu_all['sol'], border=NA)
```